---
title: "F24- Spatial_Pre-processing"
author: "Niels"
date: "`r Sys.Date()`"
output: pdf_document
---

This script preprocesses the cultural heritage (monuments) and gadm data.
The necessary data includes anlægsdata, which contains a data base over cultural heritage sites in Denmark and can be accessed from here: https://www.kulturarv.dk/fundogfortidsminder/

It also uses municipality data from the gadm data website, which can be found here: https://gadm.org/about.html

And finally it uses postal code data accesses through the DAWA (Danish Address Web API) portal here: https://dawadocs.dataforsyningen.dk/dok/api/postnummer#

# Loading Packages
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
pacman::p_load(sf, tidyverse, jsonlite)
```

# Preprocessing the monuments data
```{r}
# Load the cultural monuments data
monuments <- read_sf("../in/raw_data/anlaeg_all_25832.shp")

# Load the municipality data
municipality <- read_sf("../in/raw_data/gadm41_DNK_2.shp")

# Filter the municipalities to include only Copenhagen
copenhagen <- municipality %>% 
  filter(NAME_2 == "København")

# Select only the relevant columns. it is very cluttered rn
monuments <- monuments %>% 
  select(systemnr, stednavn, stednavnsb, anlaegsbet, datering, fra_aar, til_aar, geometry)

# Ensure the CRS matches
monuments <- st_transform(monuments, crs = st_crs(copenhagen))

# Filter monuments so that it is only those within Copenhagen
monuments <- monuments[st_intersects(monuments, copenhagen, sparse = FALSE), ]

# Extract coordinates from the geometry in monuments column and making a longitude and latitude column
coordinates <- st_coordinates(monuments)
monuments$longitude <- coordinates[, 1]
monuments$latitude <- coordinates[, 2]
```


# Sanity check to make sure that all the monuments are within copenhagen
```{r}
plot(st_geometry(copenhagen), col = 'lightblue', main = "Monuments within Copenhagen")
plot(st_geometry(monuments), col = 'red', add = TRUE)
```

# processing the postal codes, and adding postal code to each monument
```{r}
# Load the postal code data
postal_codes <- fromJSON("../in/raw_data/postnumre.json")

# Convert postal codes data to a data frame and clean it
postal_codes_df <- as.data.frame(postal_codes) %>% 
  select(postal_code = nr, area_name = navn, bbox) %>% # selecting only the used columns and renaming them. bbox is the coordinates
  unnest_wider(bbox, names_sep = "_") # unnesting the coordinates

# Create a function to find the postal code based on coordinates. It checks the coordinates of the monuments to see if any fall within the bbox (polygons) of a postal code
find_postal_code <- function(lon, lat, postal_codes_df) {
  for (i in 1:nrow(postal_codes_df)) { # goes through each row in the postal codes
    bbox <- as.numeric(postal_codes_df[i, c("bbox_1", "bbox_2", "bbox_3", "bbox_4")]) # the bbox is made up of xmin (1), xmax (3), ymin(2), ymax(4).
    if (lon >= bbox[1] && lon <= bbox[3] && lat >= bbox[2] && lat <= bbox[4]) {
      return(postal_codes_df$postal_code[i]) # if the coordinates of the monuments is within the bbox, it is assigned to that postal code
    }
  }
  return(NA)
}

# Add postal code to the monuments data
monuments$postal_code <- mapply(find_postal_code, monuments$longitude, monuments$latitude, MoreArgs = list(postal_codes_df = postal_codes_df)) # apply the function to make a new row in the monuments df

# sanity check to see if it works
head(monuments) #
```

# PLOTTING THE POSTAL CODE BOUNDARIES to make sure that it works
```{r}
# convert bounding boxes to sf objects for plotting
postal_codes_sf <- postal_codes_df %>%
  rowwise() %>%
  mutate(geometry = st_sfc(st_polygon(list(matrix(c(bbox_1, bbox_2, bbox_3, bbox_2, bbox_3, bbox_4, bbox_1, bbox_4, bbox_1, bbox_2), ncol = 2, byrow = TRUE))))) %>%
  st_as_sf(crs = st_crs(copenhagen))

# Plot the bounding boxes and monuments for sanity check
plot(st_geometry(postal_codes_sf), col = NA, border = 'grey', main = "Postal Codes and Monuments")
plot(st_geometry(monuments), col = 'red', add = TRUE)
```


# further processing the monument data
```{r}
print(unique(monuments$anlaegsbet)) # getting total list of different types of monuments in the data set
# The monuments df contains A LOT of different heritage sites. Only picking a subset here (mostly buildings as they are difficult to move in case of a flood)
list_monuments <- c("Bygning", "Fæstningsanlæg", "Kirkegård", "Klosteranlæg", 
                    "Gravkapel", "Hovedbygning", "Kirke", "Herregårdsanlæg/Slot", 
                    "Embedsbolig", "Rådhus", "Fattiggård", "Vindmølle", 
                    "Skanse", "Borg/Voldsted", "Vandmølle", "Kapel", 
                    "Ødekirke", "Administrationsbygning", "Genforeningssten", 
                    "Fængsel", "Fontæne", "Kampplads", "Staldbygning", 
                    "Kaserne", "Mindesmærke", "Monument", "Teater", 
                    "Badehus", "Hospital", "Kulturminde", "Ruin", 
                    "Historisk anlæg")

# Only select the above chosen monuments
monuments <- monuments %>% 
  filter(anlaegsbet %in% list_monuments)
```


```{r}
# Check for missing values in the df
sum(is.na(monuments$geometry))
sum(is.na(monuments$anlaegsbet))
# ther are none

# Some of the entries in the monuments df contains multiple entries of the same building (mostly, if the building was present for more than one time period, it is recorded twice). This does not make sense for our analysis, therefore we are removing the oldest of the duplicated entries
sum(duplicated(monuments$systemnr))

monuments <- monuments %>%
  arrange(systemnr, desc(fra_aar)) %>%
  distinct(systemnr, .keep_all = TRUE)
```


```{r}
# Saving the data for the processing script
st_write(copenhagen, "../in/preprocessed_data/Copenhagen_muni.shp")
st_write(monuments, "../in/preprocessed_data/Monuments.shp")
```

