---
title: "Spatial_Analytics"
author: "Niels"
date: "2024-06-01"
output: pdf_document
---

## R Markdown

This script analyses the data. Besides what is needed in the preprocessing, it also uses the Danmarks Højdemodel (DHM), which is a digital model of the height of the terrain compared to the sea level. This version Havvand på Land of the DHM further simulates how different areas are flooded if the water level rises. It can be access either through WMS by using their API or as GEOTIFF both from here (https://dataforsyningen.dk/data/2695)

# Setup

# loading packages
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
pacman::p_load(sf, raster, terra, tidyverse)
```

```{r}
# Load the preprocessed data of the copenhagen muniplicity and the monuments within the borders
monuments <- read_sf("../in/preprocessed_data/Monuments.shp")
copenhagen <- read_sf("../in/preprocessed_data/Copenhagen_muni.shp")

# loading the DHM. The GEOTIFF version is grided i0 10x10km, therefore multiple grids were needed to be downloaded. So its a lot of files.
tif_files <- list.files(path = "../in/raw_data/Hoejdemodel", pattern = "*.tif", full.names = TRUE)
rasters <- lapply(tif_files, raster) # converting the tiffs into a raster
merged_raster <- do.call(merge, rasters) # merging all of the DHM rasters

# Check and transform CRS if needed
raster_crs <- crs(merged_raster)
copenhagen_crs <- st_crs(copenhagen)

# making sure the two data sets use the same CRS. If not then the cph muniplicity df is converted to use the DHM CRS (as doing the opposite way would take a loooot of processing)
if (!identical(as.character(copenhagen_crs$proj4string), as.character(raster_crs@projargs))) {
  copenhagen <- st_transform(copenhagen, crs = as.character(raster_crs))
}

# Crop the merged raster to the extent of Copenhagen. Even though I only downloaded the grids that outline Copenhagen, there is still redundant data (as the grids were 10x10km)
copenhagen_bbox <- st_bbox(copenhagen)
copenhagen_extent <- extent(copenhagen_bbox$xmin, copenhagen_bbox$xmax, copenhagen_bbox$ymin, copenhagen_bbox$ymax)
cropped_raster <- crop(merged_raster, copenhagen_extent)

# Extract elevation data for the monuments. It takes each monument and finds the corresponding height in the raster from that position. Then it assigns that height to the monument
monument_elevations <- extract(cropped_raster, monuments)

# adding them to the monuments
monuments$elevation <- monument_elevations

# Define water levels for floodiung height
water_levels <- c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6)

# Function that assigns a flooding level to each of the monuments. It takes the height of the monument, and if that height is equal or less than a given water level, then that becomes the flood level. As the model only goes to 6 meters, if the require water level to flood the monument is higher than 6m, then it will be safe 
determine_flood_level <- function(elevation, levels) {
  flood_level <- max(levels[levels <= elevation], na.rm = TRUE)
  if (is.na(flood_level)) return(7) else return(flood_level)
}

# Apply the function to all of the monuments
monuments$flood_level <- sapply(monuments$elevation, determine_flood_level, levels = water_levels)

# save the monuments with the added flooding levels
st_write(monuments, "../in/preprocessed_data/Monuments_with_flood_levels.shp")
```


















