---
title: "F24-Sptial_Plotting"
author: "Niels"
date: "`r Sys.Date()`"
output: pdf_document
---
Plotting part of the analysis. For this part, we are going to use the same data sets as before. However, we are goping to load the DHM model through WMS instead of using the TIFF files. I had some issues using the TIFF files, so going with the WMS approach instead
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
pacman::p_load(sf, leaflet, tidyverse, htmlwidgets)
```


```{r}
# Loading the preprocessed data
monuments <- read_sf("../in/preprocessed_data/Monuments_with_flood_levels.shp")
copenhagen <- read_sf("../in/preprocessed_data/Copenhagen_muni.shp")

# Verify the CRS of the two data sets
st_crs(monuments)
st_crs(copenhagen)
```
#  quick tally to find out how many momuments in total
```{r}
size_sum(monuments)
```




# defining the depth model and getting the layers from the API
```{r}
# defining the api for the wms, which is downloaded from Dataforsyningen. You can make your own token by making an account on their website.
wms_url <- "https://api.dataforsyningen.dk/dhm?service=WMS&request=GetCapabilities&token=1baf42948629b2b5f3b9995b9cdf0d61" # probably need to make a new token to retrieve the layers (later)

# defnining the different layers of height and the corresponding layers from the WMS
water_levels <- list(
  "Water level of 0.5m" = "dhm_gummistoevleindeks_havvand_50cm",
  "Water level of 1m" = "dhm_gummistoevleindeks_havvand_100cm",
  "Water level of 1.5m" = "dhm_gummistoevleindeks_havvand_150cm",
  "Water level of 2m" = "dhm_gummistoevleindeks_havvand_200cm",
  "Water level of 2.5m" = "dhm_gummistoevleindeks_havvand_250cm",
  "Water level of 3m" = "dhm_gummistoevleindeks_havvand_300cm",
  "Water level of 3.5m" = "dhm_gummistoevleindeks_havvand_350cm",
  "Water level of 4m" = "dhm_gummistoevleindeks_havvand_400cm",
  "Water level of 4.5m" = "dhm_gummistoevleindeks_havvand_450cm",
  "Water level of 5m" = "dhm_gummistoevleindeks_havvand_500cm",
  "Water level of 5.5m" = "dhm_gummistoevleindeks_havvand_550cm",
  "Water level of 6m" = "dhm_gummistoevleindeks_havvand_600cm"
)
```

# Leaflet plot. For some reason something in the preprocessing is removing all the vowels from the variable names
```{r}
leaflet_map <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = copenhagen, color = "blue", weight = 1, fillOpacity = 0.1, group = "Copenhagen") %>%
  addCircleMarkers(data = monuments, 
                   color = "yellow", 
                   radius = 3, 
                   label = ~anlgsbt, 
                   popup = ~paste("<strong>Type:</strong>", anlgsbt, "<br><strong>ID:</strong>", systmnr, "<br><strong>Flood Level:</strong>", fld_lvl, "m"), 
                   group = "Monuments")

# adding the layers from the WMS
for (level in names(water_levels)) {
  leaflet_map <- leaflet_map %>%
    addWMSTiles(
      baseUrl = wms_url,
      layers = water_levels[[level]],
      options = WMSTileOptions(format = "image/png", transparent = TRUE),
      group = paste("Water level of", sub("Water level of ", "", level))
    )
}
# adding the controls
leaflet_map <- leaflet_map %>%
  addLayersControl(
    overlayGroups = c("Copenhagen", "Monuments", paste("Water level of", sub("Water level of ", "", names(water_levels)))),
    options = layersControlOptions(collapsed = FALSE),
    position = 'topright'
  ) %>%
  hideGroup(paste("Water level of", sub("Water level of ", "", names(water_levels))))

# defining legends
legend_colors <- c("red", "blue")
legend_labels <- c("Flooded", "Semi-flooded")

#adding legands
leaflet_map <- leaflet_map %>%
  addLegend(position = "bottomleft",
            colors = legend_colors,
            labels = legend_labels,
            title = "Flood Levels",
            opacity = 1)

# show the map
leaflet_map

# save the map to display on my repo readme
saveWidget(leaflet_map, "../out/leaflet_map.html", selfcontained = TRUE) 
```


# prepping for plots showing the total number of monuments and the total number of monuments at flood level by postal code
```{r}
# Summarize the number of monuments at flood level by postal code and risk level, excluding risk level 6
summary_by_postal_code <- monuments %>%
  filter(fld_lvl != 6) %>%
  group_by(pstl_cd, fld_lvl) %>%
  summarize(count = n(), .groups = 'drop')

# Summarize the total number of monuments at risk by flood level, excluding risk level 6
summary_by_water_level <- monuments %>%
  filter(fld_lvl != 6) %>%
  group_by(fld_lvl) %>%
  summarize(count = n(), .groups = 'drop')

```

# Plots for showing the total number of monuments at flood risk overall and by postal code
```{r}
# Plot the number of monuments at flood level by postal code and flood level
ggplot(summary_by_postal_code, aes(x = pstl_cd, y = count, fill = fld_lvl)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Cultural Heritage Sites at Risk of Flooding by Postal Code and Water Level",
       x = "Postal Code",
       y = "Number of Monuments",
       fill = "Water Level (Meters)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the total number of monuments at risk by flood level
ggplot(summary_by_water_level, aes(x = fld_lvl, y = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Cultural Heritage Sites at Risk of Flooding by Water Level",
       x = "Water Level",
       y = "Number of Monuments",
       fill = "Water Level (Meters)") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(0, 6, by = 0.5))
```

# Create table for summary of the results
```{r}
summary_table <- monuments %>%
  group_by(fld_lvl) %>%
  summarise(number_of_monuments = n()) %>%
  arrange(fld_lvl)

summary_table <- summary_table %>%
  mutate(fld_lvl = factor(fld_lvl, levels = c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6)))

print(summary_table)
```

